Estimation of Intracortical Connectivity in a Dynamical Neural Field Model

### 2010-03-16 ltx/BrainIDE.tex
some more updates to the paper

Even though we will model  $v_0$ as being homogeneous across the cortical sheet we will express it as being spatially dependant allowing for generality when constructing the state-space representation in the next section. The parameters in equation~\ref{ActivationFunction} are detailed in Table 1. By combining Eqs.~\ref{SpikesToPotential} and \ref{RateBasedInteractions} we get the spatiotemporal model
\begin{equation}\label{FullDoubleIntModel}
v\left(r,t\right) = \int\limits_{-\infty}^t h\left(t - t'\right) \left(\int\limits_\Omega  w\left(r,r'\right)f\left( v\left( r',\bar t \right)\right)dr' + e\left(r,t\right)\right) dt'
\end{equation}
To arrive at the final form of the model we shall state the synaptic response kernel as a Green's function
\begin{equation}\label{GreensFuncDef}
Lh\left( t \right) = \delta \left( t \right),
\end{equation}
where $L=\frac{d}{dt} + \alpha$ is a temporal differential operator and $\delta(t)$ is the Dirac-delta function. This provides the most general form
\begin{equation}\label{FinalForm1}
dv\left( {r,t} \right) + \alpha v\left( r,t \right)dt = \int_\Omega  {w\left( r,r' \right)f\left( {v\left( r',\bar t \right)} \right)dr'dt}  +de\left( r,t \right).
\end{equation}

\section{Discrete Time and Space Model}
We begin the derivation of the state-space model with the continuous time, continuous space IDE neural field model. 
Spatio-temporal interactions can be described by the continuous space continuous time stochastic differential equation
\begin{equation}\label{differential eq}	
dv\left( r,t \right) =  \left( -\alpha v\left( r,t \right) + \int_\Omega  {w\left( r,r' \right)f\left( {v\left( r',t \right)} \right)dr'} \right) dt  + d e\left( r,t \right).
\end{equation}
To simplify the notation, but without loss of generality, \ref{FinalForm1} can be rewritten as
\begin{equation}\label{differential simplified}
	dv\left( r,t \right) = q\left(v(r,t) \right) dt  + d e\left( r,t \right)
\end{equation}
A one-step Euler method was used to approximate the nonlinear stochastic IDE to a discrete time form. 
\begin{equation}\label{euler}
	v\left(r,t\right) =  v(r,t_0) + \int_{t_0}^{t} q\left(v(r,\tau) \right) d\tau  + \int_{t_0}^{t} d e\left(r,\tau\right)
\end{equation}
Using an Euler approximation,~\ref{euler} can be approximated at each sampling time by
\begin{equation}\label{euler approx 1}
	v\left(r,t_{k+1} \right) =  v(r,t_k) + q\left(v(r,t_k)\right) (t_{k+1} - t_k) + (e(r,t_{k+1}) - e(r,t_k))
\end{equation}
which can be rewritten as
\begin{equation}\label{euler approx 2}
	v\left(r,t_k + \delta \right) =  v(r,t_k) + \delta q\left(v(r,t_k)\right) + e'(r,t_k)
\end{equation}
where $e'(r,t_k)$ is a zero mean Gaussian random variable with covariance $\delta\Sigma_e$. (Note that, because of the way the noise is defined in \ref{differential eq} it is not $\delta^2$ but $\delta$). Expanding $q\left(v(r,t_k)\right)$ we have
\begin{equation}\label{discrete time model}
	v\left(r,t_k + \delta \right) =  v(r,t_k) + \delta  \left( -\alpha v\left( r,t_k \right) + \int_\Omega  {w\left( r,r' \right)f\left( {v\left( r',t_k \right)} \right)dr'} \right) + e'(r,t_k)
\end{equation}
which can be simplified to
\begin{equation}\label{discrete time model2}
	v\left(r,t_k + \delta \right) =  (1 - \delta \alpha) v(r,t_k) + \delta \int_\Omega  {w\left( r,r' \right)f\left( {v\left( r',t_k \right)} \right)dr'} + e'(r,t_k)
\end{equation}
Now we define the spatial aspect of the model on a regular square $i,j$ grid of neural masses, where the spatial step size $\Delta \mathbf{r}_i = \Delta \mathbf{r}_j = \Delta $ giving
\begin{equation}\label{discrete space}
	v\left(r_{ij},t_k + \delta \right) =  (1 - \delta \alpha)v(r_{ij},t_k) + \delta \Delta^2 \sum_{i=-I}^{i=I}{\sum_{j=-J}^{j=J}{\tilde{w}\left( r_{ij},r_{ij}' \right)f\left( v\left( r_{ij}',t_k \right) \right)} }+ e'(r_{i,j},t_k).
\end{equation}
where $\tilde{w}(r_{ij},r_{ij}')$ discretized connectivity kernel. In Eq~\ref{discrete space} we have used periodic boundary conditions to alleviate problems associated with edge effects. The connectivity kernel is in the form of a modified Mexican hat, and is described by the sum of three weighted Gaussian functions as
\begin{equation}
	\tilde{w}(r_{ij},r_{ij}') = \mathbf{\theta}^T e^{\frac{\left(\left(r_i'-r_i\right)\Delta\right)^2}{(2\mathbf{\sigma})^2} + \frac{\left(\left(r_j'-r_j\right)\Delta\right)^2}{(2\mathbf{\sigma})^2}}
\end{equation}
where
\begin{equation}
 \mathbf{\theta}=\begin{bmatrix} \mathbf{\theta}_{1}\\\mathbf{\theta}_{2}\\\mathbf \theta_{3}\end{bmatrix},\quad \mathbf{\sigma}=\begin{bmatrix}\mathbf{\sigma}_1\\\mathbf{\sigma}_2\\ \mathbf{\sigma}_3\end{bmatrix}.
\end{equation}.
The first Gaussian function (parametrized by $\mathbf{\theta}_1, \mathbf{\sigma}_1$) describes local excitation, the second describes lateral inhibition and the third describes medium range excitation.

>>>>

Even though we will model  $v_0$ as being homogeneous across the cortical sheet we will express it as being spatially dependant allowing for generality when constructing the state-space representation in the next section. The parameters in equation~\ref{ActivationFunction} are detailed in Table 1. By combining Eqs.~\ref{SpikesToPotential} and \ref{RateBasedInteractions} we get the spatiotemporal model
\begin{equation}\label{FullDoubleIntModel}
v\left(r,t\right) = \int\limits_{-\infty}^t h\left(t - t'\right) \left(\int\limits_\Omega  w\left(r,r'\right)f\left( v\left( r',\bar t \right)\right)dr' + e\left(r,t\right)\right) dt'
\end{equation}
The rearrange the model into an intergro-differential we take the Laplace transform of~\ref{FullDoubleIntModel} giving
\begin{equation}
	V(s) = H(s)G(s).
\end{equation}
Since $H(s) = (s+\alpha)^{-1}$ we can write
\begin{equation}
	(s+\alpha)V(s) = G(s).
\end{equation}
% To arrive at the final form of the model we shall state the synaptic response kernel as a Green's function
% \begin{equation}\label{GreensFuncDef}
% Lh\left( t \right) = \delta \left( t \right),
% \end{equation}
% where $L=\frac{d}{dt} + \alpha$ is a temporal differential operator and $\delta(t)$ is the Dirac-delta function. 
Now taking the inverse Laplace transform provides the final form
\begin{equation}\label{FinalForm1}
dv\left( r,t \right) + \alpha v\left( r,t \right)dt = \int_\Omega  {w\left( r,r' \right)f\left( {v\left( r',\bar t \right)} \right)dr'dt}  +de\left( r,t \right).
\end{equation}

\section{Discrete Time and Space Model}
We begin the derivation of the state-space model with the continuous time, continuous space IDE neural field model. 
Spatio-temporal interactions can be described by the continuous space continuous time stochastic differential equation
\begin{equation}\label{differential eq}	
dv\left( r,t \right) =  \left( -\alpha v\left( r,t \right) + \int_\Omega  {w\left( r,r' \right)f\left( {v\left( r',t \right)} \right)dr'} \right) dt  + d e\left( r,t \right).
\end{equation}
To simplify the notation, but without loss of generality, \ref{FinalForm1} can be rewritten as
\begin{equation}\label{differential simplified}
	dv\left( r,t \right) = q\left(v\left(r,t\right) \right) dt  + d e\left( r,t \right)
\end{equation}
where $q\left(v\left(r,t\right)\right)$ is the term insides the brackets in Eq.~\ref{differential eq}. A one-step Euler method was used to approximate the nonlinear stochastic IDE to a discrete time form. 
\begin{equation}\label{euler}
	v\left(r,t\right) =  v(r,t_0) + \int_{t_0}^{t} q\left(v(r,\tau) \right) d\tau  + \int_{t_0}^{t} d e\left(r,\tau\right)
\end{equation}
Using an Euler approximation,~\ref{euler} can be approximated at each sampling time by
\begin{equation}\label{euler approx 1}
	v\left(r,t_{k+1} \right) =  v(r,t_k) + q\left(v(r,t_k)\right) (t_{k+1} - t_k) + (e(r,t_{k+1}) - e(r,t_k))
\end{equation}
which can be rewritten as
\begin{equation}\label{euler approx 2}
	v\left(r,t_k + \delta \right) =  v(r,t_k) + \delta q\left(v(r,t_k)\right) + e'(r,t_k)
\end{equation}
where $e'(r,t_k)$ is a zero mean Gaussian random variable with covariance $\delta\Sigma_e$. (Note that, because of the way the noise is defined in \ref{differential eq} it is not $\delta^2$ but $\delta$). Expanding $q\left(v(r,t_k)\right)$ we have
\begin{equation}\label{discrete time model}
	v\left(r,t_k + \delta \right) =  v(r,t_k) + \delta  \left( -\alpha v\left( r,t_k \right) + \int_\Omega  {w\left( r,r' \right)f\left( {v\left( r',t_k \right)} \right)dr'} \right) + e'(r,t_k)
\end{equation}
which can be simplified to
\begin{equation}\label{discrete time model2}
	v\left(r,t_k + \delta \right) =  (1 - \delta \alpha) v(r,t_k) + \delta \int_\Omega  {w\left( r,r' \right)f\left( {v\left( r',t_k \right)} \right)dr'} + e'(r,t_k)
\end{equation}
Now we define the spatial aspect of the model on a regular square $i,j$ grid of neural masses, where the spatial step size $\Delta \mathbf{r}_i = \Delta \mathbf{r}_j = \Delta $ giving
\begin{equation}\label{discrete space}
	v\left(r_{ij},t_k + \delta \right) =  (1 - \delta \alpha)v(r_{ij},t_k) + \delta \Delta^2 \sum_{i=-I}^{i=I}{\sum_{j=-J}^{j=J}{\tilde{w}\left( r_{ij},r_{ij}' \right)f\left( v\left( r_{ij}',t_k \right) \right)} }+ e'(r_{i,j},t_k).
\end{equation}
where $\tilde{w}(r_{ij},r_{ij}')$ discretized connectivity kernel. In Eq~\ref{discrete space} we have used periodic boundary conditions to alleviate problems associated with edge effects. The connectivity kernel is in the form of a modified Mexican hat, and is described by the sum of three weighted Gaussian functions as
\begin{equation}
	\tilde{w}(r_{ij},r_{ij}') = \mathbf{\theta}^T e^{\frac{\left(\left(r_i'-r_i\right)\Delta\right)^2}{(2\mathbf{\sigma})^2} + \frac{\left(\left(r_j'-r_j\right)\Delta\right)^2}{(2\mathbf{\sigma})^2}}
\end{equation}
where
\begin{equation}
 \mathbf{\theta}=\begin{bmatrix} \mathbf{\theta}_{1}\\\mathbf{\theta}_{2}\\\mathbf \theta_{3}\end{bmatrix},\quad \mathbf{\sigma}=\begin{bmatrix}\mathbf{\sigma}_1\\\mathbf{\sigma}_2\\ \mathbf{\sigma}_3\end{bmatrix}.
\end{equation}.
The first Gaussian function (parametrized by $\mathbf{\theta}_1, \mathbf{\sigma}_1$) describes local excitation, the second describes lateral inhibition and the third describes medium range excitation.
